---
title: "myexam"
output: html_document
date: "2024-11-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

```         
A sports scientist collects strength data from two groups of students. The first group is a group of athletes who might be expected to be strong, and the second group is a control group of nonathletic students.
The first group has strength ratings of: 
c(127.8, 133.6, 129.7, 130.4, 128.7, 132.6, 129.6, 132.6)
And the second group has strength ratings of:
c(137.9, 107.3, 126.9, 113.3, 127.9, 104.1)
Is there evidence that the first group is actually stronger than the second?
(a) State a sensible null hypothesis
(b) State the precise definition of p-value and explain what “more extreme” means in this
context
(c) Is a one-sided or two-sided test needed? justify
(d) Perform a student t-test using R and interpret
```

```{r sports}
group1 <- c(127.8, 133.6, 129.7, 130.4, 128.7, 132.6, 129.6, 132.6)
group2 <- c(137.9, 107.3, 126.9, 113.3, 127.9, 104.1)
```

### a. null hypothesis

The null hypothesis is that group1 of athletes and group2 of non-athletes have the same strength on average.

H~0~: mean(group1) == mean(group2)

H~1~: mean(group1) \> mean(group2)

### b. p-value

The p-value is the probability, if the null is true, of obtaining the observation or an observation more extreme. More extreme being that group1 is stronger than group 2 of non-athletes.

### c. one-sided

This is one-side, because we expect the athletes to be stronger than the non-athletes.

### d. student-test

```{r}
t.test(group1, group2, alternative = "greater")
```

The `p-value` is less than 5%. The result is statistically significant, and we have strong evidence that gorup 1 of athletes is stronger than group 2 of non-athletes.

## Question 2

```         
Random variable X is drawn from a normal distribution with mean 6.71 and standard devi-
ation 2.31.
(a) Calculate the probability of X being less than 3.99.
(b) What is the probability of X exceeding 6.11?
(c) What is the probability of X lying between 6.72 and 9.08?
(d) Verify your answers to parts (a) and (b) above using numerical sampling.
(e) (Harder) verify your answers to part (c) above using numerical sampling.
```

### a. The probbility of X being less than 3.99

```{r}
mean <- 6.71
sd <-2.31
pnorm(3.99, mean = mean, sd = sd)
```

The probability of X being less than 3.99 is `0.1195002`

### b. the probability of X exceeding 6.11

```{r}
1 - pnorm(6.11, mean = mean, sd = sd)
```

The probability of X exceeding 6.11 is `0.6024679`.

### c. the probability of X lying between 6.72 and 9.08

```{r}
pnorm(9.08, mean = mean, sd = sd) - pnorm(6.72, mean = mean, sd = sd)
```

The probability of X lying between 6.72 and 9.08 is `0.3458211`.

### d. Verify your answers to parts (a) and (b) above using numerical sampling.

```{r}
sample <- rnorm(1e6, mean = mean, sd = sd)
table(sample < 3.99) / 1e6
table(sample > 6.11) / 1e6

```

Using numerical sampling, the probabilities using numerical sampling are very close (to three decimal places) to the calculated values of `0.1195002`, and `0.6024679`.

### (Harder) verify your answers to part (c) above using numerical sampling.

```{r}
sample <- rnorm(1e6, mean = mean, sd = sd)
table(sample > 6.72 & sample < 9.08) / 1e6
```

The probability using numerical sampling is the same (to three decimal places) to the calculated value of `0.3458211`.

## Question 3

```         
A structural scientist is studying an engine that has 8 cylinders in it. He takes an engine
and subjects it to a standard operational test. The cylinders are subject to stresses and as
time goes by sometimes develop cracks. The probability of a cylinder developing a crack
(this is a tiny crack, visible only with specialized equipment) is 0.318.
(a) If the cylinders’ status is independent, what distribution specifies the total number of
cracked cylinders?
(b) Using dbinom() or otherwise, what is the probability that exactly two cylinders are
cracked?
(c) What is the probability that four or more cylinders are cracked?
(d) A new engine is subject to the operational test but using a different type of lubricating
oil, which is claimed to reduce stress in the engine. The scientist finds that only one
cylinder is cracked. State a sensible null hypothesis, and report its p-value. Interpret
your finding in a way that a busy engineer, who is not a statistician, could understand
(e) [harder] What is the maximum value of p which will ensure that only 10% of engines
suffer four or more cracked cylinders? (you need only quote an approximate solution;
2 decimal places is sufficient for maximum credit).
```

### a. what distribution specifies the total number of cracked cylinders

The binomial distribution.

### b. Using dbinom() or otherwise, what is the probability that exactly two cylinders are cracked

```{r}
dbinom(2, 8, prob = 0.318)
```

The probability is `0.2849171`.

### c. What is the probability that four or more cylinders are cracked?

**WRONG** I USED 1 - pbinom(4,....), it should be 1 - pbinom(3,....)

```{r}
1 - pbinom(3, 8, 0.318)
```

another way to check:

```{r}
sum(dbinom(4:8, 8, 0.318))
```
